---
title: "Homework3"
author: "Brian Yu"
date: "October 12, 2016"
output: html_document
---

```{r}
sessionInfo()
```
#### Use R to do each of the following. Use R code instructions that are a general as possible, and also as efficient as possible.

##### 1. Using the `testdata` data vector you created for Lecture 1 practice problems:
  45.4  44.2  36.8  35.1  39.0  60.0  47.4  41.1  45.8  35.6

1a. Use logical referencing to calculate the standard deviation of only those values that are less than 45.
```{r}
testdata <- c(45.4, 44.2, 36.8, 35.1, 39.0, 60.0, 47.4, 41.1, 45.8, 35.6)
# Display contests in testdata
testdata
sortHL.testdata <- sort(testdata, decreasing = TRUE)
sortHL.testdata
stdevLT45.testdata <- sd(sortHL.testdata[-c(1, 2, 3, 4)])
stdevLT45.testdata
```

1b. Write an R command that will determine how many vector entries are less than 45. Display the command and its output
```{r}
sum(testdata < 45)
```

1c. Write an R command that will determine how many vector entries are greater than 40 but less than 55. Display the command and its output.
```{r}
sum(testdata > 40 & testdata < 55)
```

1d. Write an R command that will calculate what proportion of the data vector has values exceeding 40.
```{r}
length(which(testdata > 40)) / length(testdata)
```


##### 2. This exercise uses a file called `cigsales.txt` that contains state-level data regarding cigarette sales and other variables. The data set is available from the Unit Overview page for this unit. Use `read.table` to import this data set as a data frame in R.

2a. The variable `black` indicates the percentage of a given state that is African-American. Using logical referencing, select from this data frame only those states that have over a 15% African-American population. What states get selected?
```{r}
# Import data set as data frame
cigdata = read.table("cigsales.txt", header = TRUE)
cigdata # Displaying data
cigdata.black.GT15 <- subset(cigdata, black > 15.0)
cigdata.black.GT15 # States with greater than 15% African-American population
```

2b. Extract the variable `price` from this data frame, and place it into a vector of its own called `price.vec`.
```{r}
price.vec <- (cigdata$price)
class(price.vec) # Verifying it's a vector
price.vec # Display vector price.vec
```

2c. Use logical referencing to create two separate vectors in R: A vector called poor containing only the income values that fall below the median income value for all the states; and a vector called rich containing only the income values that fall above the median income value for all the states.
```{r}
income.vec <- (cigdata$income) # Extract out income column from cigdata data frame
income.vec # Display vector income.vec
poor <- income.vec[income.vec < median(income.vec)] # Create vector poor
poor # Display vector poor
rich <- income.vec[income.vec > median(income.vec)] # Create vector rich
rich # Display vector rich
```

##### 3. Run the R code examples below and write a brief statement about what each example does

3a. Example 1:
```{r}
trim <- function(x, lower = 0.0, upper = 1.0) {
  indicies <- x >= lower & x <= upper
  return(x[indicies])
}

x <- sort(rnorm(20)) # Create new vector x
x # Display vector x
y <- trim(x) # Create new vector y utilizing trim function
y # Display vector y

    # This function trims off the values of the argument x that are above or below the
    # arguments lower (0.0) and upper limit (1.0)
```

3b. Example 2:
```{r}
trim2 <- function(x, lower = 0.0, upper = 1.0) x[x >= lower & x <= upper]

y2 <- trim2(x)
y2

    # This function does the same action as Example 1 without a return statement. The
    # function returns the value of the last expression if there is no return.
```

3c. Example 3:
```{r}
sumdice <- function(n) {
  k <- sample(1:6, size=n, replace = TRUE)
  return(sum(k))
}

sumdice(1) # The sumdice function, n is the number of dice to roll and add
```

##### 4. On average, how many dots would you expect to see when you roll a fair six-sided die? How can you use the function `sumdice()` to get a reliable estimate of the average ("expected") number of dots?
```{r}
mean(replicate(100000, sumdice(1)))
    # On average, approximately 3.5 dots is expected. To yield this result, I used
    # sumdice() as a subfunction of the replicate function with 100,000 rolls, which was 
    # a subfunction of the mean function.
```

##### 5. Examine the following piece of R code. What is the difference between the variable x and the variable y? (Note: I am not asking for a number. I am asking for the difference in how they function within the code below. Write your answer in sentence form).
```{r}
x <- 42
fred <- function(y) {
  x <- y
  return(y + x)
}
fred(13)
x
    # Inside the function, x is a local variable that is defined to be the value of y.
    # However, outside the function,x is unchanged. For example, if you use the function
    # fred(13), this means the local variable x is 13 since it is defined to be the value
    # of y, which is 13, thus the return statement is (13 + 13) = 26. However, the x
    # variable outside the function is unchanged at 42.
```

##### 6. Create an R function `diff()` that calculates the absolute difference between the mean and median of a sample of values stored in a vector. Then, run your function on the `price.vec` vector from part 2b.
```{r}
diff <- function(z) {
  k <- abs(median(z) - mean(z))
  return(k)
}
diff(price.vec) # Running diff() function on vector price.vec
```

##### 7. Here is a very simple loop in R. Run this code and tell me what it does:
```{r}
for (i in 1:20) {
  print(i + 3)
}
    # This for loop takes an argument i, which is a vector from 1 through 20 and 
    # prints out the results by adding +3 to every value from 1 through 20.
```

