---
title: "LiveSessionUnit11Assignment"
author: "Brian Yu"
date: "November 20, 2016"
output: 
  html_document:
    keep_md: true
---
# Overview
For Live Session Unit 11 Assignment, I was assigned to analye the ukcars dataset.The deliverable for this assignment is a link to R Markdown file on GitHub.

```{r, message=FALSE, warning=FALSE}
library(fpp) #fpp package must be installed first
```

##### a) Plot the time series. Can you identify seasonal fluctuations and/or a trend?
```{r}
data(ukcars)
plot(ukcars)
```
There seems to be a seasonality compenent to this time series

##### b) Use a classical decomposition to calculate the trend-cycle and seasonal indicies.
```{r}
fitd <- decompose(ukcars)
trend_indicies <- fitd$trend
seasonal_indicies <- fitd$seasonal
plot(fitd)
```

##### c) Do the results support the graphical interpretation from part (a)?
```{r}
head(trend_indicies, 113)
head(seasonal_indicies, 113)
```
By decomposing the time series, a definitive seasonal fluctuation trend. From Q1 to Q2 of a year, the trend decrease. From Q3 to Q4, the trend increases.

##### d) Compute and plot the seasonally adjusted data.
```{r}
fit_season_adj <- seasadj(fitd)
plot(fit_season_adj)
```

##### e) Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
```{r}
ts <- ukcars
ts[50] = ts[50] + 500
fitd2 <- decompose(ts)
fit_season_adj2 <- seasadj(fitd2)
plot(fit_season_adj2)
```

At index 50, we added 500 to the observation value, and it causes a spike in the plot at index 50.

##### f) Does it make any difference if the outlier is near the end rather than in the middle of the time series?
```{r}
ts2 <- ts
ts[100] = ts[100] + 500
fitd2 <- decompose(ts)
fit_season_adj3 <- seasadj(fitd2)
plot(fit_season_adj3)
```

From observation of the graph, it doesn't seem to make a difference if the outlier is near the end rather than in the middle of the time series

##### g) Use STL to decompose the series.
```{r}
fitstl <- stl(ukcars, s.window = 5)
plot(fitstl)
```
